Shaping Anti-Corruption Campaigns in Nigeria

### 2024-12-22 paper/paper.qmd
final edits
final typos fixed!

---
title: "Shaping Anti-Corruption Campaigns in Nigeria"
subtitle: "Youth Populations, Urban Areas and Moderately Educated Citizens Are Important for Anti-Corruption Campaigns "
author: 
  - Fatimah Yunusa
thanks: "Code and data are available at:https://github.com/fatimahsy/Nigeria_Democracy-"
date: today
date-format: long
abstract: "Peceptions of corruption in the Nigerian presidency vary significantly across demogrpahic groups. Using Afrobarometer survey data and Bayesisan ordinal logistic regression, this study looks at how education, gender, age, and urban versus rural residence influence these perceptions. The results show that urban residents and younger individuals are more likely to perceive high corruption, while rural populations and those with higher education show views that are moderate. These results show the need for targeted anti-corruption campaigns to adress demographic differences and imporve governance and accountability in Nigeria"
format: pdf
toc: true
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

@@

Corruption is a big challenge to governance and development in Nigeria as a result of its ability to undermine public trust and economic growth. As the United Nations Office on Drugs and Crime (UNODC) stated, "corruption affects the lives of ordinary Nigerian, impeding development and eroding public trust in institutions [@corruptionsurvey2024]. This paper looks into the role of demographic factors-education,gender,age, and urban versus rural residence-in shaping perceptions of corruption in the Nigerian presidency.

@@

Our primary measurement task is to capture public opinion towards
economic, political and socioeconomic issues and translate it into
actionable lintel. The Afrobarometer survey translates these real world
perceptions into quantifiable data using survey instruments.The process
begins with identifying relevant phenomena in the Nigerian
socio-political and economic context, such as public attitudes toward
democracy, demand for governance accountability, and perceived economic
conditions. These phenomena are operationalized into survey questions
that are clear, culturally sensitive, and tailored to elicit meaningful
responses [@Afrobarometer2022codebook].

@@

Each survey question acts as a data proxy for these real-world
dynamics, reducing them to quantifiable variables while keeping
their conceptual integrity.This dataset shows a unique moment in Nigeria's socio-political
landscape, showing how citizens perceive democracy and governance in
the context of ongoing economic and political challenges. Alternative
datasets, such as Nigeria’s national statistics or World Bank survey
data, could provide economic indicators but do not have the granularity of
Afrobarometer’s public opinion data.

The data for this study was downloaded, cleaned, analyzed, modeled, and
visualized using R [@citeR], a statistical programming language.[@afrobarometer2023] served as the primary data source, providing importantsurvey data for analysis.***Telling Stories with Data*** [@tellingstories] was referenced for its methodologies. The
following packages were used:

@@

@tbl-cleaned-data-preview represents the first six cleaned rows of the
dataset, showing the important variables used in the analysis. The
variables were constructed from specific survey questions in the
Afrobarometer Round 9 Nigeria Survey [@afrobarometer2023], which
captures several socioeconomic sentiments within the country.The column
`corruption_presidency` captures perceptions of corruption within the
office of the presidency, derived from responses to the question, *“How
many of the following people do you think are involved in corruption, or
haven’t you heard enough about them to say: the president and officials
in his office?”* The variable education reflects respondents' highest
level of education completed, gathered from the question, *“What is the
highest level of education you have completed?”* The gender variable
indicates the respondent's gender, collected through direct demographic
inquiry. The column age_group categorizes respondents into age brackets
based on the question,* “How old are you?”* The urban_rural variable
distinguishes between urban and rural respondents, as recorded by
interviewers based on the primary sampling unit (PSU) during data
collection [@survey_manual_2022]. Lastly, region specifies the
geographical region of each respondent, identified using a unique code
for Nigerian states. This table not only demonstrates the diversity of
the dataset but also highlights the breadth of demographic and
governance-related variables included for analysis.

@@

```{r}
#| label: fig-education-vs-corruption
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| fig.cap: "Relationship between Education Levels and corruption percetpions in Nigeria"

@@

The heatmap shows that respondents who completed secondary school or
some secondary schooling dominate the High Corruption and Some
Corruption categories. Respondents with higher education levels
(university and post-graduate) are distributed more evenly across the
corruption perception categories, with fewer concentrated in any
specific category. Conversely, respondents with no formal schooling or
only informal education are sparsely represented across all
categories.There is a notable concentration of corruption perceptions
among respondents with mid-level education (secondary schooling), while
those with higher education tend to have more varied perceptions. These
findings suggest a potential link between educational attainment and
experiences or attitudes toward corruption, particularly among those
with secondary-level education.

@@

The 25–34 and 35–44 age groups have the highest counts in the High
Corruption and Some Corruption categories, suggesting that middle-aged
respondents are more likely to perceive corruption as significant. In
contrast, younger respondents (18–24) and older respondents (55+) report
lower levels of corruption perception in these categories.In the All Corrupt category, the 25–34, 35–44, and 55+ age groups dominate, while the 18–24 and 45–54 groups have lower counts. The No
Corruption category has low counts across all age groups, indicating
that very few respondents, regardless of age, perceive the absence of
corruption.

@@

The relationship between respondents' residential location (rural vs.
urban) and their perceptions of corruption is shown in
@fig-ruralurban-versus-corruption Each bar represents the proportion of
respondents within rural and urban groups who hold different corruption
perceptions.The chart shows that rural and urban respondents share similar
patterns of corruption perception. The majority of respondents in both
groups fall under the "Some Corruption" and "High Corruption" categories,
with "Some Corruption" being slightly more prevalent. The "All Corrupt"
category is the third most common perception in both groups, while the
"No Corruption" category is consistently the least represented.

@@

The threshold parameters ($\alpha_j$) separate the ordered categories of
the outcome variable, while the Dirichlet prior for $\pi_{ij}$ assumes
no strong prior preference for any specific corruption category.The use of these
priors provides flexibility for the model while preventing overfitting. The 
posterior distributions for the coefficients and thresholds, along with 
diagnostic checks for model convergence and predictive accuracy, can be found in
the Appendix.

@@

The analysis revealed trends in corruption perceptions in
different demographics. Our results are summarized in @tbl-modelresults. 
This presents the estimated coefficients and standard
errors for each predictor included in the model. The dependent variable
is the perception of corruption in the presidency, categorized into
ordered levels. The table includes predictors such as education level,
gender, urban or rural location, and agre group, which are modeled to
explain variations in corruption perceptions. Each predictor's
coefficient represents its effect on the log odds of perceiving
corruption in the presidency at a specific level relative to lower
levels, holding other variables constant.

Higher education levels(e.g., university or post-graduate
qualifications) are associated with positive effects on corruption
perceptions, indicating a greater likelihood of perceived higher levels
of corruption. Being female shows a small positive association with
higher corruption perceptions. Residing in rural areas is negatively
associated with corruption perceptions, suggesting lower perceptions of
corruption compared to urban counterparts. Age groups show minimal
variation in their association with corruption perceptions, with no
significant deviations observed among different age brackets.

@@

## Tareting Education Levels in Anti-Corruption Campaigns

@@

This paper is focused on looking at trends in the corruption perception
between different demographics and seeing how these differences can be
used to maximize anti corruption campaigns and policy in
Nigeria.From this paper, we find different actionable strategies for
improving the effectiveness of anti-corruption campaigns in Nigeria.
First, targeting rural and urban populations differently might be
effective. Urban residents, who have greater access to investigative
media and public debates, are more likely to perceive high corruption.
Campaigns in urban areas should therefore focus on digital platforms
like social media and mobile apps, as well as partnerships with
investigative journalists and public accountability platforms to engage
this group effectively. In rural areas, where reliance on patronage
systems and limited media access shape perceptions, grassroots
communication strategies like community radio, local town halls, and
collaborations with trusted community leaders are better suited to
disseminate anti-corruption messaging.

Education also emerges as an important determinant of corruption perceptions.
Secondary school graduates exhibit the highest levels of awareness, this
suggests that civic education campaigns should focus on this demographic
to harness their increased engagement. Integrating anti-corruption
education into school curricula could early awareness and long-term
attitudinal changes.Youth engagement is very important, given that
younger age groups, particularly those aged 25–34, report the highest
levels of dissatisfaction with corruption. Social media campaigns
tailored to platforms popular with youth, together with interactive town
hall discussions and youth-led grassroots initiatives, could channel
this demographic's energy into organized anti-corruption efforts.

@@

While this paper tells us a lot about the demographic distributions of
corruption perception about the presidency in Nigeria, there are a few
limitations that have to be considered. The survey data relies on
self-reported perception, which can be influenced by social desirability
bias which refers to the situation where respondents' endorse more
favorable responses in order to enhance their own
self-presentation[@bias2022].For example, high-profile corruption
scandals or electoral promises happening close to survey periods may
distort responses. The data does not account for regional differences
within these urban rural classifications, which could change the
understanding of corruption perceptions further.The last limitation is
with the bayesian model, because while the Bayesian model accounts for
uncertainty, it does not address potential unobserved variables, like
political affiliations or socioeconomic status, which may affect the
relationships we say.

@@

## Aforbarometer Methodology Overview and Evaluation {#sec-survey-appendix}

@@

Afrobarometer is a  pan-African, non-partisan survey research network established to measure public attitudes on democracy, governance, economic reform, civil society, and quality of like in Africa. It relies predominantly on in-person interviews using a standard questionnaire. The project covers up to 40 countries in Africa. The information collected is intended to be used by various parties including policymakers, activists,teachers and scholars. The surveys are conducted using a standardized questionnaire tailored to specific countries through a process called "indigenization", ensuring relevance to local contexts while keeping comparability across nations [@survey_manual_2022].

@@

### Smpling Approach

@@

- **Exclusion of Institutional Populations**: Individuals in prisons, military barracks, and similar institutions are excluded, potentially omitting perspectives from disenfranchised groups [@survey_manual_2022].
- **Gender Quotas**: The enforcement of equal gender representation ensures that women's voices are equally captured, addressing historical disparities in research participation [@survey_manual_2022].
- **Regional and Security Constraints**: Some areas, particularly conflict zones, may be inaccessible, leading to under representation in survey findings.This is particularly the cause in North-Eastern Nigeria which as a result of criminal groups like Boko Haram, has made it virtually impossible to access that region safely.

@@

The ethical considerations underpinning Afrobarometer’s methodology ensure the credibility and validity of its findings while safeguarding respondent dignity. However, addressing remaining gaps—such as the exclusion of certain populations and challenges in conflict zones—would further enhance the ethical rigor of corruption perception research in Africa. By adhering to established ethical frameworks and remaining responsive to emerging challenges, studies like this can continue to give important information while maintaining the highest ethical standards.To do my part in ensuring ethical standards are upheld, i have included a datasheet for the Afrobarometer dataset.

@@

This figure provides an overview of demographic variables and corruption perceptions from the Afrobarometer Nigeria Round 9 survey. Rural residents constitute a larger share of the sample, aligning with Nigeria's predominantly rural demographic landscape. Corruption perception data shows that "High Corruption" and "Some Corruption" dominate public opinion. The education level distribution highlights the prevalence of basic education, with the majority of respondents having completed secondary school. Finally, the age distribution indicates that the 25–34 age group is the most represented cohort.

![Demographic Distributions of the Afrobarometer Data](combined_plot.png){#fig-demo width="110%"}

>>>>

---
title: "Shaping Anti-Corruption Campaigns in Nigeria"
subtitle: "Youth Populations, Urban Areas and Moderately Educated Citizens Are Important for Anti-Corruption Campaigns "
author: 
  - Fatimah Yunusa
thanks: "Code and data are available at: [https://github.com/fatimahsy/Nigeria_Democracy-/tree/main](https://github.com/fatimahsy/Nigeria_Democracy-/tree/main)"
date: today
date-format: long
abstract: "Perceptions of corruption in the Nigerian presidency vary significantly across demographic groups. Using Afrobarometer survey data and Bayesian ordinal logistic regression, this study looks at how education, gender, age, and urban versus rural residence influence these perceptions. The results show that urban residents and younger individuals are more likely to perceive high corruption, while rural populations and those with higher education show views that are moderate. These results show the need for targeted anti-corruption campaigns to address demographic differences and improve governance and accountability in Nigeria"
format: pdf
toc: true
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

@@

Corruption is a big challenge to governance and development in Nigeria as a result of its ability to undermine public trust and economic growth. As the United Nations Office on Drugs and Crime (UNODC) stated, "corruption affects the lives of ordinary Nigerian, impeding development and eroding public trust in institutions [@corruptionsurvey2024]. This paper looks into the role of demographic factors-education, gender, age, and urban versus rural residence-in shaping perceptions of corruption in the Nigerian presidency.

@@

Our primary measurement task is to capture public opinion towards
economic, political and socioeconomic issues and translate it into
actionable lintel. The Afrobarometer survey translates these real world
perceptions into quantifiable data using survey instruments. The process
begins with identifying relevant phenomena in the Nigerian
socio-political and economic context, such as public attitudes toward
democracy, demand for governance accountability, and perceived economic
conditions. These phenomena are operationalized into survey questions
that are clear, culturally sensitive, and tailored to elicit meaningful
responses [@Afrobarometer2022codebook].

@@

Each survey question acts as a data proxy for these real-world
dynamics, reducing them to quantifiable variables while keeping
their conceptual integrity. This dataset shows a unique moment in Nigeria's socio-political
landscape, showing how citizens perceive democracy and governance in
the context of ongoing economic and political challenges. Alternative
datasets, such as Nigeria’s national statistics or World Bank survey
data, could provide economic indicators but do not have the granularity of
Afrobarometer’s public opinion data.

The data for this study was downloaded, cleaned, analyzed, modeled, and
visualized using R [@citeR], a statistical programming language.[@afrobarometer2023] served as the primary data source, providing important survey data for analysis.***Telling Stories with Data*** [@tellingstories] was referenced for its methodologies. The
following packages were used:

@@

@tbl-cleaned-data-preview represents the first six cleaned rows of the
dataset, showing the important variables used in the analysis. The
variables were constructed from specific survey questions in the
Afrobarometer Round 9 Nigeria Survey [@afrobarometer2023], which
captures several socioeconomic sentiments within the country. The column
`corruption_presidency` captures perceptions of corruption within the
office of the presidency, derived from responses to the question, *“How
many of the following people do you think are involved in corruption, or
haven’t you heard enough about them to say: the president and officials
in his office?”* The variable education reflects respondents' highest
level of education completed, gathered from the question, *“What is the
highest level of education you have completed?”* The gender variable
indicates the respondent's gender, collected through direct demographic
inquiry. The column age_group categorizes respondents into age brackets
based on the question,* “How old are you?”* The urban_rural variable
distinguishes between urban and rural respondents, as recorded by
interviewers based on the primary sampling unit (PSU) during data
collection [@survey_manual_2022]. Lastly, region specifies the
geographical region of each respondent, identified using a unique code
for Nigerian states. This table not only demonstrates the diversity of
the dataset but also highlights the breadth of demographic and
governance-related variables included for analysis.

@@

```{r}
#| label: fig-education-vs-corruption
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| fig.cap: "Relationship between Education Levels and corruption perceptions in Nigeria"

@@

The heatmap shows that respondents who completed secondary school or
some secondary schooling dominate the High Corruption and Some
Corruption categories. Respondents with higher education levels
(university and post-graduate) are distributed more evenly across the
corruption perception categories, with fewer concentrated in any
specific category. Conversely, respondents with no formal schooling or
only informal education are sparsely represented across all
categories. There is a notable concentration of corruption perceptions
among respondents with mid-level education (secondary schooling), while
those with higher education tend to have more varied perceptions. These
findings suggest a potential link between educational attainment and
experiences or attitudes toward corruption, particularly among those
with secondary-level education.

@@

The 25–34 and 35–44 age groups have the highest counts in the High
Corruption and Some Corruption categories, suggesting that middle-aged
respondents are more likely to perceive corruption as significant. In
contrast, younger respondents (18–24) and older respondents (55+) report
lower levels of corruption perception in these categories. In the All Corrupt category, the 25–34, 35–44, and 55+ age groups dominate, while the 18–24 and 45–54 groups have lower counts. The No
Corruption category has low counts across all age groups, indicating
that very few respondents, regardless of age, perceive the absence of
corruption.

@@

The relationship between respondents' residential location (rural vs.
urban) and their perceptions of corruption is shown in
@fig-ruralurban-versus-corruption Each bar represents the proportion of
respondents within rural and urban groups who hold different corruption
perceptions. The chart shows that rural and urban respondents share similar
patterns of corruption perception. The majority of respondents in both
groups fall under the "Some Corruption" and "High Corruption" categories,
with "Some Corruption" being slightly more prevalent. The "All Corrupt"
category is the third most common perception in both groups, while the
"No Corruption" category is consistently the least represented.

@@

The threshold parameters ($\alpha_j$) separate the ordered categories of
the outcome variable, while the Dirichlet prior for $\pi_{ij}$ assumes
no strong prior preference for any specific corruption category. The use of these
priors provides flexibility for the model while preventing overfitting. The 
posterior distributions for the coefficients and thresholds, along with 
diagnostic checks for model convergence and predictive accuracy, can be found in
the Appendix.

@@

The analysis revealed trends in corruption perceptions in
different demographics. Our results are summarized in @tbl-modelresults. 
This presents the estimated coefficients and standard
errors for each predictor included in the model. The dependent variable
is the perception of corruption in the presidency, categorized into
ordered levels. The table includes predictors such as education level,
gender, urban or rural location, and age group, which are modeled to
explain variations in corruption perceptions. Each predictor's
coefficient represents its effect on the log odds of perceiving
corruption in the presidency at a specific level relative to lower
levels, holding other variables constant.

Higher education levels (e.g., university or post-graduate
qualifications) are associated with positive effects on corruption
perceptions, indicating a greater likelihood of perceived higher levels
of corruption. Being female shows a small positive association with
higher corruption perceptions. Residing in rural areas is negatively
associated with corruption perceptions, suggesting lower perceptions of
corruption compared to urban counterparts. Age groups show minimal
variation in their association with corruption perceptions, with no
significant deviations observed among different age brackets.

@@

## Targeting Education Levels in Anti-Corruption Campaigns

@@

This paper is focused on looking at trends in the corruption perception
between different demographics and seeing how these differences can be
used to maximize anti corruption campaigns and policy in
Nigeria. From this paper, we find different actionable strategies for
improving the effectiveness of anti-corruption campaigns in Nigeria.
First, targeting rural and urban populations differently might be
effective. Urban residents, who have greater access to investigative
media and public debates, are more likely to perceive high corruption.
Campaigns in urban areas should therefore focus on digital platforms
like social media and mobile apps, as well as partnerships with
investigative journalists and public accountability platforms to engage
this group effectively. In rural areas, where reliance on patronage
systems and limited media access shape perceptions, grassroots
communication strategies like community radio, local town halls, and
collaborations with trusted community leaders are better suited to
disseminate anti-corruption messaging.

Education also emerges as an important determinant of corruption perceptions.
Secondary school graduates exhibit the highest levels of awareness, this
suggests that civic education campaigns should focus on this demographic
to harness their increased engagement. Integrating anti-corruption
education into school curricula could early awareness and long-term
attitudinal changes. Youth engagement is very important, given that
younger age groups, particularly those aged 25–34, report the highest
levels of dissatisfaction with corruption. Social media campaigns
tailored to platforms popular with youth, together with interactive town
hall discussions and youth-led grassroots initiatives, could channel
this demographic's energy into organized anti-corruption efforts.

@@

While this paper tells us a lot about the demographic distributions of
corruption perception about the presidency in Nigeria, there are a few
limitations that have to be considered. The survey data relies on
self-reported perception, which can be influenced by social desirability
bias which refers to the situation where respondents' endorse more
favorable responses in order to enhance their own
self-presentation[@bias2022].For example, high-profile corruption
scandals or electoral promises happening close to survey periods may
distort responses. The data does not account for regional differences
within these urban rural classifications, which could change the
understanding of corruption perceptions further. The last limitation is
with the Bayesian model, because while the Bayesian model accounts for
uncertainty, it does not address potential unobserved variables, like
political affiliations or socioeconomic status, which may affect the
relationships we say.

@@

## Afrobarometer Methodology Overview and Evaluation {#sec-survey-appendix}

@@

Afrobarometer is a  pan-African, non-partisan survey research network established to measure public attitudes on democracy, governance, economic reform, civil society, and quality of like in Africa. It relies predominantly on in-person interviews using a standard questionnaire. The project covers up to 40 countries in Africa. The information collected is intended to be used by various parties including policymakers, activists, teachers and scholars. The surveys are conducted using a standardized questionnaire tailored to specific countries through a process called "indigenization", ensuring relevance to local contexts while keeping comparability across nations [@survey_manual_2022].

@@

### Sampling Approach

@@

- **Exclusion of Institutional Populations**: Individuals in prisons, military barracks, and similar institutions are excluded, potentially omitting perspectives from disenfranchised groups [@survey_manual_2022].
- **Gender Quotas**: The enforcement of equal gender representation ensures that women's voices are equally captured, addressing historical disparities in research participation [@survey_manual_2022].
- **Regional and Security Constraints**: Some areas, particularly conflict zones, may be inaccessible, leading to under representation in survey findings. This is particularly the cause in North-Eastern Nigeria which as a result of criminal groups like Boko Haram, has made it virtually impossible to access that region safely.

@@

The ethical considerations underpinning Afrobarometer’s methodology ensure the credibility and validity of its findings while safeguarding respondent dignity. However, addressing remaining gaps—such as the exclusion of certain populations and challenges in conflict zones—would further enhance the ethical rigor of corruption perception research in Africa. By adhering to established ethical frameworks and remaining responsive to emerging challenges, studies like this can continue to give important information while maintaining the highest ethical standards. To do my part in ensuring ethical standards are upheld, i have included a datasheet for the Afrobarometer dataset.

@@

This figure provides an overview of demographic variables and corruption perceptions from the Afrobarometer Nigeria Round 9 survey. Rural residents constitute a larger share of the sample, aligning with Nigeria's predominantly rural demographic landscape. Corruption perception data shows that "High Corruption" and "Some Corruption" dominate public opinion. The education level distribution highlights the prevalence of basic education, with the majority of respondents having completed secondary school. Finally, the age distribution indicates that the 25–34 age group is the most represented cohort.
![Demographic Distributions of the Afrobarometer Data](combined_plot.png){#fig-demo width="110%"}
